---
title: "MAST90139 Assignment 3"
author: "Haonan Zhong 867492 (Wed 14:15 Qiuyi Li)"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Question 1a
The model fitted in the $\mathtt{R}$ output is a multi-categorical logistic model, which use status as the nominal categorical response variable, and year as the predictor. The response variable status with three levels follows a multinomial distribution M(371, $\boldsymbol{\pi}$), where $\boldsymbol{\pi}$ is a vector that contains probability of being in each level.

The model fitted to the data is
$$
log\frac{P(mild)}{P(normal)} = -4.2917 + 0.0836\times year
$$
$$
log\frac{P(severe)}{P(normal)} = -5.0598 + 0.1093\times year
$$

### Question 1b
0.1093 is the coefficient for year when status = severe, therefore the intepretation is given by: for every additional year spent working at the coal face the odds of a miner's pneu-monoconiosis status is severe will be increased by a factor of $e^{0.1093}$ = 1.115497.

$$
\frac{P(severe)}{P(normal)} = e^{-5.0598 + 0.1093\times year}
$$
\newline
$$
Odds\:ratio = \frac{e^{-5.0598 + 0.1093\times 10}}{e^{-5.0598 + 0.1093\times 0}} = e^{0.1093\times 10} = 2.98321
$$
Therefore, the 95% confidence interval for the odds ratio of severe status versus normal status for every 10 more years spent working at the coal face is given by the following:
$$
(e^{10\times(1.093-1.96\times 0.0165)}, e^{10\times(1.093+1.96\times 0.0165)}) = (2.158903, 4.122254)
$$

### Question 1c
Estimate the pneumonoconiosis status probabilities for a miner who has spent 25 years working at the coal face, using normal level as the pivot.
$$
\frac{P(severe)}{P(normal)} = e^{-5.0598 + 0.1093\times 25}
$$
\newline
$$
\frac{P(mild)}{P(normal)} = e^{-4.2917 + 0.0836\times 25}
$$
Therefore,
$$
P(normal) = \frac{1}{1+e^{-4.2917 + 0.0836\times 25}+e^{-5.0598 + 0.1093\times 25}}
$$
Hence,
$$
P(mild) = P(normal)\times e^{-4.2917 + 0.0836\times 25}
$$
$$
P(severe) = P(normal)\times e^{-5.0598 + 0.1093\times 25}
$$
```{r}
odds_mild = exp(-4.2917 + 0.0836*25)
odds_severe = exp(-5.0598 + 0.1093*25)
p_normal = 1/(1 + odds_mild + odds_severe)
p_mild = p_normal * odds_mild
p_severe = p_normal * odds_severe
cbind(p_normal, p_mild, p_severe)
```
Therefore, the probability of normal, mild and severe is 0.8276955, 0.0915555 and 0.08074898 respectively. And given they sum up to one, which means our estimated probabilities are valid.

### Question 2a
The model fitted in question 2 is a cumulative logistic model, response variable is status as an ordinal variable and covariate is year. And the model can be written as follow
$$
log\frac{P(Status=Normal)}{P(Status>Normal)}=3.9558-0.0959\times year
$$
\newline
$$
log\frac{P(Status < Severe)}{P(Status = Severe)}=4.8690-0.0959\times year
$$

### Question 2b
The interpretation of the coefficient estimate 0.0959 is given by: the odds of status moving from normal to mild/severe, or moving from normal/mild to severe increases by a factor of $e^{0.0959}$ = 1.100649 for every additional year a coal miner spent working at the coal face.

$$
\frac{P(Status=Normal)}{P(Status>Normal)} = e^{3.9558-0.0959\times year}
$$
$$
\frac{P(Status>Normal)}{P(Status=Normal)}=\frac{1}{e^{3.9558-0.0959\times year}}
$$
$$
Odds\:ratio=\frac{e^{3.9558-0.0959\times 0}}{e^{3.9558-0.0959\times 10}}=\frac{1}{e^{-0.0959\times 10}}=e^{0.0959\times 10}=2.6091
$$
The estimated odds ratio of non-normal status versus normal status for every 10 more years spent working at the coal face is 2.6091.Furthermore, the 95% confidence interval for the odds ratio of severe status versus normal status for every 10 more years spent working at the coal face is given by the following:
$$
(e^{10\times(0.0959 - 1.96\times 0.01194)}, e^{10\times(0.0959 + 1.96\times 0.01194)}) = (2.064682, 3.297036)
$$

### Question 2c
Estimate the pneumonoconiosis status probabilities for a miner who has spent 25 years working at the coal face
$$
\frac{P(Status>Normal)}{P(Status=Normal)}=\frac{P(Mild) + P(Severe)}{P(Normal)} = \frac{1-P(Normal)}{P(Normal)} = \frac{1}{e^{3.9558 - 0.0959\times 25}}
$$
Solving the above equation, we have
$$
P(Normal) = \frac{1}{e^{3.9558 - 0.0959\times 25}+1} = 0.8261
$$

$$
\frac{P(Status<Severe)}{P(Status=Severe)} = \frac{P(Normal)+P(Mild)}{P(Severe)} = \frac{1-P(Severe)}{P(Severe)}=\frac{1}{e^{4.8690-0.0959\times 25}}
$$
Solving the above equation, we have
$$
P(Severe) = \frac{1}{e^{4.8690-0.0959\times 25}+1} = 0.0779
$$
And since probability sum up to one, therefore
$$
P(Mild) = 1 - P(Normal) - P(Severe) = 1 - 0.8261 - 0.0779 = 0.0960
$$
Therefore, the estimated probability are P(Normal) = 0.8261, P(Mild) = 0.0960, and P(Severe) = 0.0779.











